Functional Pearl: Short and Mechanized Logical Relation for Dependent Type Theories

### 2024-02-28 paper/paper.tex
add future work to conclusion

\section{Conclusion}
\label{sec:conclusion}
In this work, we present a short and mechanized proof of consistency through
proof by logical relations for a dependently typed language with a full
universe hierarchy, large eliminations, an intensional identity type, and
dependent eliminators. We show the extensibility of our approach by proving
the existence of $\beta\eta$-normal forms with only small and mechanical
changes to our proof development. Our Coq mechanization leverages existing Coq
libraries for reasoning about metatheory and for general purpose automation,
allowing us to significantly reduce the verbosity typically associated with
mechanized proofs. The result is a declarative proof style that is usually
only available on pen and paper. Our development shows that proofs by logical
relations for dependent types are accessible and do not
require months of effort to implement. We hope our proof can inspire
researchers to more frequently mechanize results, such as consistency and
normalization, for their dependent type theories.

>>>>

\section{Future work and Conclusions}
\label{sec:conclusion}
In this work, we present a short and mechanized proof of consistency through
proof by logical relations for a dependently typed language with a full
universe hierarchy, large eliminations, an intensional identity type, and
dependent eliminators.  We show the extensibility of our approach by proving
the existence of $\beta\eta$-normal forms with only small and mechanical
changes to our proof development. Related work gives us confidence that we
could extend our logical relation to include features such as full inductive
datatypes, irrelevant arguments, and type-directed conversion; though it is
not clear how much of the brevity of the development we will be able to
maintain.  Furthermore, we hope that mechanized logical relations proofs will
eventually grow to include other features found in dependent type theories,
such as impredicative universes, universe polymorphism, and cumulativity.

Our Coq mechanization leverages existing Coq libraries for reasoning about
metatheory and for general purpose automation, allowing us to significantly
reduce the verbosity typically associated with mechanized proofs. The result
is a declarative proof style that is usually only available on pen and
paper. Our development shows that proofs by logical relations for dependent
types are accessible and do not require months of effort to implement.
We hope our proof can inspire researchers to more frequently mechanize
results, such as consistency and normalization, for their dependent type
theories. By design, the goals of our work are modest.
